---
title: "R Notebook"
output: html_notebook
---
Proof-of-concept for OSM feature extraction (and preprocessing) in R.

References: (https://dominicroye.github.io/en/2018/accessing-openstreetmap-data-with-r/) 

```{r, init}
library(tidyverse)
library(osmdata)
library(sf)
library(ggmap)
```

We need to ask OSM servers what features are available, so that we can filter out. The `available_features()` function returns a list of available OSM features that have different tags. Feature names and possible tags are kept up-to date at the [OSM wiki](https://wiki.openstreetmap.org/wiki/Map_features) (be prepared, is rather longish and well documented!).

In our case, we need "anything road", so let's ask... 

```{r get_available_features}
# have a look at what's available
available_features()
```
Whoa.

Lotsa stuff here. Reading [the wiki](https://wiki.openstreetmap.org/wiki/Highways), we can be safe asking for "`highway`" 

```{r get_available_tags}
# have a look at highways
available_tags("highway")
```
Seems good.
Looks like we need "`highway`" indeed. We also have to define some spatial boundaries, since we just want to deal with -- say -- a slice of Northern Italy. Fix that bounding box issue, theh back to tamper with OSM.

:question: :construction: Note that in `osmdata` one explicits bounding boxes by place names, whereas real mapmakers use real bounding boxes. So much for the Muggles. Well'fix that later.

```{r make_bb}
# we have an AOI polygon in BASE, I used the boundaries of Piedmont and Lombardy, plus a 10 km buffer, then reprojected from 23032 into 4326.
AOI <- st_read(dsn="BASE", layer="AOI-NW-ITA_4326")

# tis is the standard osmdata mechanism, what if we have an extent?
bb <- getbb("Piemonte")
background.map <- get_map(bb, maptype = "toner-background")
ggmap(background.map)

```
So far, so good, we can (albeit in a silly way) hook up a part of the world. Back to `highway`, now.

```{r get_highways}
# note the use of tidyverse piping
# note 2 set a timeout value high enough to opq
q <- getbb("Piemonte") %>% opq(timeout=50000) %>% add_osm_feature("highway")
roads <- osmdata_sf(q) # this takes ages

# well, try a plot
ggmap(background.map) + geom_sf(data=roads$osm_points,
                                inherit.aes = FALSE,
          colour = "#238443",
          fill = "#004529",
          alpha = .5,
          size = 4,
          shape = 21)+
  labs(x = "", y = "")

```
